Digital enhancement of the ``Paolo Orsi'' museum: a Google Street View 360° pilot project tour

### 2015-12-16 source/20/EAGLE2016_BONACINI - Pilot project at Paolo Orsi Museum.tex
added bst files in each folder, started work on bibliography through Zotero, added CNR paper

\title{}
\author{}
\date{2015-12-11}
\begin{document}
Digital enhancement of the “Paolo Orsi” museum: a Google Street View 360° pilot project tour \newline

Elisa Bonacini

Humanities Department, University of Catania; IEMEST (Istituto Euro Mediterraneo di Scienze e Tecnologie), Palermo,
Italy

Email: e\_bonacini@hotmail.com; elisa.bonacini@unict.it

\newline
Abstract

\ \ \ \ \ \ The aim of this paper is to present the pilot project in progress at the “Paolo Orsi” Archaeological Museum
(Syracuse, Italy). Thanks to a free partnership with Google Business Photos, we have managed to map the entire museum
for a online 360° tour on Google Street View. A dozen archaeological finds have been selected for 360° virtual tours,
provided with descriptive sheets. Among them the beautiful ‘inscription of Nassiane’, from the Catacombs of San
Giovanni in Syracuse, has been selected.

\bigskip

Keywords

Google, Sicily, Sicilian Cultural Heritage, Virtual Museum, Virtual tour, Digital Heritage.

\bigskip

\bigskip

1.1. The “Paolo Orsi” Archaeological Museum

The “Paolo Orsi” Regional Archaeological Museum of Syracuse, together with the “Antonino Salinas” Regional
Archaeological Museum in Palermo, is the most important Sicilian archaeological museum and it is one of the most
important and richest archaeological museums in Italy.

The National Archaeological Museum of Syracuse was born by a royal decree in 1878, known as “national” for its
collections’ importance and size. Well-placed inside the historical palace in the Cathedral Square on Ortigia island,
it was directed by Paolo Orsi from 1895 to 1934.

@@

\bigskip

1.2. The project: reasons and birth

@@

Sicily has the highest number of UNESCO heritage sites (7/51 in total\footnote{1997: Valley of Temples in Agrigento;
1997: Villa del Casale; 2000: Eolian Islands; 2002: Late Baroque Towns of the Val di Noto (South-Eastern Sicily); 2005:
Syracuse and the rock necropolisof Pantalica; 2013: Mount Etna; 2015: Arab-Norman sites, Palermo and the Cathedral
Churches of Cefalù and Monreale.}) and of UNESCO intangible cultural heritage (3/6\footnote{2008: Opera dei Pupi,
Sicilian puppet theatre; 2013: Mediterranean diet (transnational); 2014: Traditional agricultural practice of
cultivating the ‘vite ad alberello’ (head-trained bush vines) of the community of Pantelleria.}) in Italy and in the
world. Infact, Sicily has some of the largest and most important archaeological sites in the world: the temple of
Concordia in Agrigento, for its exceptional state of preservation has become the symbol of UNESCO itself.

Despite of this, Sicilian cultural heritage struggles to be present on Google’s platforms such as Google Art Project and
others, as it should, compared to other Italian cultural sites (Bonacini 2013; Bonacini 2014).

@@

In Street View Gallery\footnote{https://www.google.com/maps/views/streetview?gl=us.}, which now has contributed a great
number of users from all over the world, thousands are the spherical and geo located photos of Italian places. However,
tightening the selection to “Landmarks of Italy”
\footnote{https://www.google.com/maps/views/streetview/italy-highlights?gl=us.}, Sicily has only 9 spherical photos,
showing the beaches of the Aeolian Islands (7), the islands of Favignana (1) and Marettimo (1).

Among Art Project’s 605 museum collections\footnote{https://www.google.com/maps/views/streetview/art-project?gl=it.}, 47
are Italian, especially relevant to Rome, Turin, Venice and Milan. One of these, not the best artistic production, is a
Sicilian contemporary collection, relevant to the International Festival of Street Artists in Giardini Naxos (Me). No
other Sicilian museum, collection or archaeological site has been included in Art Project.

Browsing on the Street View Gallery, 21 sites are all over inscribed on the World Wonders
Project\footnote{https://www.google.com/maps/views/streetview/world-wonders-project?gl=it.}: for Italy, only Pompeii
and the historic center of Florence are inscribed in this still restricted list. Many more sites can be visited
virtually from the same project linked on Google Cultural Institute’s
website\footnote{www.google.com/intl/it/culturalinstitute/worldwonders/. }: 172 sites in the world, 22 in Italy and,
finally, 2 of them in Sicily. Until last year the Unesco site of the baroque town Val di Noto was the only one in
Sicily; currently Mount Etna, has been added.

@@

This project was born in collaboration with Mr. Gianfranco Guccione, a certified Google Business Photo photographer,
while he was working as a freelance consultant at the General Direction of the Regional Department for Cultural
Heritage and Sicilian Identity in 2014. He proposed to realize the Street View mapping of a museum and an
archaeological site in Sicily, considering the possibility of creating “augmented” virtual tours (3D virtual tours of
objects displayed in museum’s windows and virtual aerial tours, with the addition of text and audio descriptions).

The profound reason of this project consists in an effort to bridge this gap, “increasing” fruition and enhancement of
Sicilian cultural heritage.

With the agreement between the General Direction of the Regional Department for Cultural Heritage and Sicilian Identity
and the European Coordination of Google Business Photo, it was decided to choose as a sample of this project two
cultural regional institutions, the “Paolo Orsi” Museum in Syracuse and the Valley of the Temples in Agrigento, both
UNESCO sites since 2005 and 1997. The project was then structured as a research fellow project at the University of
Catania and carried out by the present writer in close collaboration with Mr. Guccione.

The first part of the project at the “Paolo Orsi” Museum, which we will discuss here, is about to be completed. The
second part at Valley of the Temples in Agrigento is going to start.

\bigskip

1.2.1. The 360° tour on Google Maps: some technical data

@@

The 360° virtual tour makes it possible browsing the entire museum and its collection, between levels connected by
arrows, placed online at the link https://goo.gl/maps/oagnd8urP1H2 (for the first level) and
https://goo.gl/maps/vrpDfuPPgwM2 (for the second level). As you can see in fig. 1.1., by clicking on the first link,
the remote user can enter the museum through the Street view panel: the 360° virtual tour begins at the entrance and
the user can browse moving with the directional arrows, between sectors and levels.

The project provided the opportunity to carry out 360° virtual tours of some exposed archaeological finds, like an
“augmented” virtual tour, certainly innovative compared to what Google’s platforms provide.

@@

\begin{flushleft}
\tablefirsthead{}
\tablehead{}
\tabletail{}
\tablelasttail{}
\begin{supertabular}{m{10.872cm}}
{\centering  \includegraphics[width=8.911cm,height=4.632cm]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img001.jpg}
\par}

~

Fig. 1.1. The virtual entrance to Archaeological Museum “Paolo Orsi” on Google Street View.

~
\\
\end{supertabular}
\end{flushleft}
As for the displaying on Google Maps of the 360° virtual tour of the objects, we must specify that Maps so far does not
support the integration of menus, captions, photos, video, info inside the Street View virtual tour technology.
Infact,it \ is only allowed navigation in 360°, ie a virtual walk.

We tried to find an answer in order to show on Maps all the information, captions, maps, levels and 360° virtual tour of
objects. Thanks to customized i-frame we made these virtual tours possible with their captions and the maps with
different levels and clickable points of interest, adding them to the existing virtual tour of the Museum, already on
Maps, through a link containing the Google mapping of the Museum. In this way the virtual tour of the Museum - made
with Google standards - is “augmented” by another virtual walk much more exhaustive, located via link on the Google
Maps board of the Museum, where you can view all these additional items.

Both the virtual tour of the museum and the virtual tour of the objects will also be placed on the Museum's website, on
the Regional Department for Cultural Heritage and Sicilian Identity portal
(http://www.regione.sicilia.it/beniculturali/museopaoloorsi/). Because of the profound principle that this non-profit
project would improve the visibility and promotion of Sicilia Cultural Heritage, the Museum’s virtual tour and its
reproductions belong to Google; Museum’s objects virtual tours, instead, belong to Paolo Orsi Museum, which is free to
reuse them.

\bigskip

1.2.2. The “Paolo Orsi” Museum 360° tour

@@

\begin{flushleft}
\tablefirsthead{}
\tablehead{}
\tabletail{}
\tablelasttail{}
\begin{supertabular}{m{10.681cm}}
{\centering  \includegraphics[width=8.98cm,height=4.697cm]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img002.jpg}
\par}

~

Fig. 1.2. 360° tour of Sector A: a couple of casts of dwarf elephants from Spinagallo cave. \\
\end{supertabular}
\end{flushleft}

\bigskip

@@

\begin{flushleft}
\tablefirsthead{}
\tablehead{}
\tabletail{}
\tablelasttail{}
\begin{supertabular}{m{10.681cm}}
{\centering  \includegraphics[width=8.89cm,height=4.641cm]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img003.jpg}
\par}

~

Fig. 1.3. 360° tour of Sector B2: the findings from Athenaion and Ionic temple in Syracuse. \\
\end{supertabular}
\end{flushleft}

\bigskip

@@

\begin{flushleft}
\tablefirsthead{}
\tablehead{}
\tabletail{}
\tablelasttail{}
\begin{supertabular}{m{10.681cm}}
{\centering  \includegraphics[width=8.708cm,height=4.521cm]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img004.jpg}
\par}

~

Fig. 1.4. 360° tour of Sector D: the Venus Landolina. \\
\end{supertabular}
\end{flushleft}
They document the multiple aspects of life in Syracuse and come from urban necropolis from III-II century B.C. Here are
masterpieces like the wonderful statue of the Venus Anadiomene, called Venus Landolina, here in Fig. 1.4.

@@

\bigskip

1.2.1.1. 360° virtual tours of archaeological selected objects

@@

[Warning: Draw object ignored]

@@

\bigskip

1.2.1.3. 360° virtual tours of \ Nassiane inscription

The last selected object we present here is a disk from Sector F: it is a marble disk, discovered by Paolo Orsi in the
Catacombs of San Giovanni in 1894 (Orsi 1895: 509-510), decorated by a wreath of laurel leaves on one side; the other
side was reused in the 4th century A.D. for the funerary inscription of Nassiane, woman who died at age 32 in God
faith.

This object, which the remote user can observe with a 360° virtual tour, has been selected both for the peculiar reuse
and because it documents the religious syncretism of the early centuries of Christianity. Re-using of architectural is
well documented in the catacombs of Syracuse (Sgarlata 2013). It was found shattered in front of a burial, distinct
from others, known as “the Saint's Tomb” (Sgarlata 2004: 40-44).

\bigskip

\begin{flushleft}
\tablefirsthead{}
\tablehead{}
\tabletail{}
\tablelasttail{}
\begin{supertabular}{m{10.681cm}}
{\centering  \includegraphics[width=10.029cm,height=5.225cm]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img005.jpg}
\par}

~

Fig. 1.5. 360° tour of the Nassiane inscription. \\
\end{supertabular}
\end{flushleft}
\ This is the Greek text:

\bigskip

chrismon \textgreek{c qrist{\textless}i{\textgreater}an~hc semn~hc >aganófronoc [>hd]`e fil'androu
}N\textgreek{assian~hc t'umbon e>isor~a|c, f'ile, k'imenon [<`wde <'h]tic semnos'unh|sin [>erize]to Phnelop'ih| }palma
chrismon \textgreek{>enj'ade k~ite [}N\textgreek{ass]ian'h, z'hsasa [>en J(e)~w|] kal~wc ka`i >am'emptwc >`eth
lb}\textgreek{', }\textgreek{m}\textgreek{~hnac i'.}

\bigskip

The inscription says: “Oh friend, you see here the tomb of Nassiane, Christian, ripe, sweet, fond of her husband, who
competed for virtue with Penelope. Here lies Nassiane, she lived well and blamelessly in God (?), 32 years and 10
months”. This inscription documents such as epigraphic formulas affected by religious syncretism (syncretism generally
means that a complex of phenomena and concepts derive from the meeting and fusion of different religious forms) between
Pagans and Christians, for which a Christian like Nassiane could have competed in life with Penelope, the hero
Odysseus’ wife and symbol of marital devotion. Its circular shape suggested that it was a table for the ritual of
refrigerium, the funeral feast that symbolically was consumed with the dead, even this syncretic practice \ dates back
to the pagan world (Sgarlata 2013, Scandurra 2014).

\bigskip

\begin{flushleft}
\tablefirsthead{}
\tablehead{}
\tabletail{}
\tablelasttail{}
\begin{supertabular}{m{10.681cm}}
{\centering  \includegraphics[width=6.205cm,height=6.174cm]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img006.jpg}
\par}

~

Fig. 1.6. The Nassiane inscription, recto (courtesy by the “Paolo Orsi” Museum). \\
\end{supertabular}
\end{flushleft}

\bigskip

1.3. Conclusions

@@

Thanks to this project we hope that Google itself could realize how the time has come to “rejuvenate” the Google Maps
Street View system, allowing enabled users to apply additional content on the maps.

However, the wide interoperability between Google software the development of new solutions and the integration of
geo-referenced results in the page results on the search engine continues unabated: the “Paolo Orsi” Museum - and with
the museum, the city of Syracuse and the whole of Sicily – will surely take advantage of this new tool for its
visibility.

\bigskip

[Warning: Draw object ignored]

\bigskip

Acknowledgement

To Mr. Gianfranco Guccione (http://www.airworks.it/) goes my heartfelt and deep thanks for having proposed to me this
project and having concluded it together, with the only aim of enhancing the Cultural Heritage in Sicily.

@@

I warmly thank all staff of the “Paolo Orsi” Regional Archaeological Museum in Syracuse for the great availability and
friendship shown to me and Mr. Guccione: Dr. Gioconda Lamagna, executive head of the Museum, Dr. Angela Maria Manenti,
head of relations with the public, all the archaeologists Dr. Anita Crispino, Dr. Agostina Musumeci and especially Dr.
Giuseppina Monterosso, tireless in supporting the project at any time and formerly member of the cabinet of Regional
Minister of Cultural Heritage and Sicilian Identity.

@@

My deepest gratitude goes to my friend Prof. Annalisa Bonica for the patience she \ has showed in correcting this
article.

This publication has been made on concession of the Regional Minister of Cultural Heritage and Sicilian Identity;
archaeological findings and plans presented here are exclusive property of the “Paolo Orsi” Regional Archaeological
Museum in Syracuse and it not allowed to be copied or duplicated \ by any means.

\bibliographystyle{sapauth-eng}
\bibliography{../../EAGLE}

>>>>

\title{Digital enhancement of the ``Paolo Orsi'' museum: a Google Street View 360° pilot project tour}
\author[catania]{Elisa Bonacini\corref{first}}
\cortext[first]{Corresponding author. Email: e_bonacini@hotmail.com; elisa.bonacini@unict.it}

\address[catania]{Humanities Department, University of Catania; IEMEST (Istituto Euro Mediterraneo di Scienze e Tecnologie), Palermo, Italy}
\date{2015-12-11}
\begin{document}

\maketitle
\begin{abstract}
The aim of this paper is to present the pilot project in progress at the ``Paolo Orsi'' Archaeological Museum
(Syracuse, Italy). Thanks to a free partnership with Google Business Photos, we have managed to map the entire museum
for a online 360° tour on Google Street View. A dozen archaeological finds have been selected for 360° virtual tours,
provided with descriptive sheets. Among them the beautiful `inscription of Nassiane', from the Catacombs of San
Giovanni in Syracuse, has been selected.
\end{abstract}
\keywords{Google, Sicily, Sicilian Cultural Heritage, Virtual Museum, Virtual tour, Digital Heritage}

\section{The ``Paolo Orsi'' Archaeological Museum}

The ``Paolo Orsi'' Regional Archaeological Museum of Syracuse, together with the ``Antonino Salinas'' Regional
Archaeological Museum in Palermo, is the most important Sicilian archaeological museum and it is one of the most
important and richest archaeological museums in Italy.

The National Archaeological Museum of Syracuse was born by a royal decree in 1878, known as ``national'' for its
collections’ importance and size. Well-placed inside the historical palace in the Cathedral Square on Ortigia island,
it was directed by Paolo Orsi from 1895 to 1934.

@@

\section{The project: reasons and birth}

@@

Sicily has the highest number of UNESCO heritage sites (7/51 in total\footnote{1997: Valley of Temples in Agrigento;
1997: Villa del Casale; 2000: Eolian Islands; 2002: Late Baroque Towns of the Val di Noto (South-Eastern Sicily); 2005:
Syracuse and the rock necropolisof Pantalica; 2013: Mount Etna; 2015: Arab-Norman sites, Palermo and the Cathedral
Churches of Cefalù and Monreale.}) and of UNESCO intangible cultural heritage (3/6\footnote{2008: Opera dei Pupi,
Sicilian puppet theatre; 2013: Mediterranean diet (transnational); 2014: Traditional agricultural practice of
cultivating the `vite ad alberello' (head-trained bush vines) of the community of Pantelleria.}) in Italy and in the
world. Infact, Sicily has some of the largest and most important archaeological sites in the world: the temple of
Concordia in Agrigento, for its exceptional state of preservation has become the symbol of UNESCO itself.

Despite of this, Sicilian cultural heritage struggles to be present on Google’s platforms such as Google Art Project and
others, as it should, compared to other Italian cultural sites \citep{Bonacini2013;Bonacini2014}).

@@

In Street View Gallery\footnote{\url{https://www.google.com/maps/views/streetview?gl=us}.}, which now has contributed a great
number of users from all over the world, thousands are the spherical and geo located photos of Italian places. However,
tightening the selection to ``Landmarks of Italy''
\footnote{\url{https://www.google.com/maps/views/streetview/italy-highlights?gl=us}.}, Sicily has only 9 spherical photos,
showing the beaches of the Aeolian Islands (7), the islands of Favignana (1) and Marettimo (1).

Among Art Project’s 605 museum collections\footnote{\url{https://www.google.com/maps/views/streetview/art-project?gl=it}.}, 47
are Italian, especially relevant to Rome, Turin, Venice and Milan. One of these, not the best artistic production, is a
Sicilian contemporary collection, relevant to the International Festival of Street Artists in Giardini Naxos (Me). No
other Sicilian museum, collection or archaeological site has been included in Art Project.

Browsing on the Street View Gallery, 21 sites are all over inscribed on the World Wonders
Project\footnote{\url{https://www.google.com/maps/views/streetview/world-wonders-project?gl=it}.}: for Italy, only Pompeii
and the historic center of Florence are inscribed in this still restricted list. Many more sites can be visited
virtually from the same project linked on Google Cultural Institute’s
website\footnote{\url{www.google.com/intl/it/culturalinstitute/worldwonders/}. }: 172 sites in the world, 22 in Italy and,
finally, 2 of them in Sicily. Until last year the Unesco site of the baroque town Val di Noto was the only one in
Sicily; currently Mount Etna, has been added.

@@

This project was born in collaboration with Mr. Gianfranco Guccione, a certified Google Business Photo photographer,
while he was working as a freelance consultant at the General Direction of the Regional Department for Cultural
Heritage and Sicilian Identity in 2014. He proposed to realize the Street View mapping of a museum and an
archaeological site in Sicily, considering the possibility of creating ``augmented'' virtual tours (3D virtual tours of
objects displayed in museum’s windows and virtual aerial tours, with the addition of text and audio descriptions).

The profound reason of this project consists in an effort to bridge this gap, ``increasing'' fruition and enhancement of
Sicilian cultural heritage.

With the agreement between the General Direction of the Regional Department for Cultural Heritage and Sicilian Identity
and the European Coordination of Google Business Photo, it was decided to choose as a sample of this project two
cultural regional institutions, the ``Paolo Orsi'' Museum in Syracuse and the Valley of the Temples in Agrigento, both
UNESCO sites since 2005 and 1997. The project was then structured as a research fellow project at the University of
Catania and carried out by the present writer in close collaboration with Mr. Guccione.

The first part of the project at the ``Paolo Orsi'' Museum, which we will discuss here, is about to be completed. The
second part at Valley of the Temples in Agrigento is going to start.

\subsection{The 360° tour on Google Maps: some technical data}

@@

The 360° virtual tour makes it possible browsing the entire museum and its collection, between levels connected by
arrows, placed online at the link \url{https://goo.gl/maps/oagnd8urP1H2} (for the first level) and
\url{https://goo.gl/maps/vrpDfuPPgwM2} (for the second level). As you can see in Fig.~\ref{fig:1}, by clicking on the first link,
the remote user can enter the museum through the Street view panel: the 360° virtual tour begins at the entrance and
the user can browse moving with the directional arrows, between sectors and levels.

The project provided the opportunity to carry out 360° virtual tours of some exposed archaeological finds, like an
``augmented'' virtual tour, certainly innovative compared to what Google’s platforms provide.

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img001.jpg}
\caption{The virtual entrance to Archaeological Museum ``Paolo Orsi'' on Google Street View. }
\label{fig:1}
\end{figure}

As for the displaying on Google Maps of the 360° virtual tour of the objects, we must specify that Maps so far does not
support the integration of menus, captions, photos, video, info inside the Street View virtual tour technology.
Infact,it is only allowed navigation in 360°, ie a virtual walk.

We tried to find an answer in order to show on Maps all the information, captions, maps, levels and 360° virtual tour of
objects. Thanks to customized i-frame we made these virtual tours possible with their captions and the maps with
different levels and clickable points of interest, adding them to the existing virtual tour of the Museum, already on
Maps, through a link containing the Google mapping of the Museum. In this way the virtual tour of the Museum - made
with Google standards - is ``augmented'' by another virtual walk much more exhaustive, located via link on the Google
Maps board of the Museum, where you can view all these additional items.

Both the virtual tour of the museum and the virtual tour of the objects will also be placed on the Museum's website, on
the Regional Department for Cultural Heritage and Sicilian Identity portal
(\url{http://www.regione.sicilia.it/beniculturali/museopaoloorsi/}). Because of the profound principle that this non-profit
project would improve the visibility and promotion of Sicilia Cultural Heritage, the Museum’s virtual tour and its
reproductions belong to Google; Museum’s objects virtual tours, instead, belong to Paolo Orsi Museum, which is free to
reuse them.

\subsection{The ``Paolo Orsi'' Museum 360° tour}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img002.jpg}
\caption{360° tour of Sector A: a couple of casts of dwarf elephants from Spinagallo cave. }
\label{fig:2}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img003.jpg}
\caption{360° tour of Sector B2: the findings from Athenaion and Ionic temple in Syracuse.}
\label{fig:3}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img004.jpg}
\caption{360° tour of Sector D: the Venus Landolina.}
\label{fig:}
\end{figure}

They document the multiple aspects of life in Syracuse and come from urban necropolis from III-II century B.C. Here are
masterpieces like the wonderful statue of the Venus Anadiomene, called Venus Landolina, here in Fig.~\ref{fig:4}.

@@

\subsubsection{360° virtual tours of archaeological selected objects}

@@

\begin{description}
\item[From Sector A:]

\begin{description}
\item[1] A foot-cup dated to the 15th century B.C., relevant to the facies of Rodì-Tindari and coming from Vallelunga, in the province of Caltanissetta. 
\end{description}

\item[From Sector B1:]

\begin{description}
\item[2] A red-figured wedding lebes, attributed to Painter of Siracusa 47099, dated to 360-340 B.C. from Lentinoi. 
\end{description}

\item[From Sector B2:]

\begin{description}
\item[3] A proto-corinthyan oinochoe, dated to 670 B.C., coming from the excavations in Piazza Duomo; 
\item[4] A plastic Corinthian vase in the form of a lion, dated 610-590 B.C., from the Garden Spain Necropolis; 
\item[5] An Attic black-figured calix-crater, by the Antimene Painter in 520 B.C., coming from the Garden Spain Necropolis; 
\item[6] An Attic black-figured Panathenaic amphora, dated to the middle of the 6th century B.C., coming from excavations in Viale Paolo Orsi; 
\item[7] A terracotta bidder statuette, dated to the 4th century B.C., coming from the sacred votive deposit in Piazza della Vittoria. 
\end{description}

\item[From Sector C:]

\begin{description}
\item[8] A red-figured bell-shaped krater, coming from Camarina, produced in the workshop of the Athenian painter Polignoto, around 440-430 B.C., decorated on the principal side with the Delphic triad (Apollo, Artemis and Latona); 
\item[9] An Attic red-figured lekythos, coming from the necropolis of Capo Soprano near Gela, dated to 470 B.C. and realized according to the manner of the London painter E342; 
\item[10] The Ephebus of Adrano, a small bronze athlete, dated to the first half of the 5th century B.C., generally thought to be a scaled-down copy of a large bronze original by the famous Greek sculptor Pythagoras. 
\end{description}

\item[From Sectors D:]

\begin{description}
\item[11] A small Hellenistic terracotta boat in the shape of a pistrix (sea monster), from the Fusco necropolis in Syracuse.

\end{description}

\item[From Sector F:]

\begin{description}
\item[12] The Nassiane inscription Fig.~\ref{fig:4}, a curious marble disk, from the Catacombs of San Giovanni.
\end{description}

\end{description}

@@

\subsubsection{360° virtual tours of Nassiane inscription}

The last selected object we present here is a disk from Sector F: it is a marble disk, discovered by Paolo Orsi in the
Catacombs of San Giovanni in 1894 \citep[509-510]{Orsi1895}, decorated by a wreath of laurel leaves on one side; the other
side was reused in the 4th century A.D. for the funerary inscription of Nassiane, woman who died at age 32 in God
faith.

This object, which the remote user can observe with a 360° virtual tour, has been selected both for the peculiar reuse
and because it documents the religious syncretism of the early centuries of Christianity. Re-using of architectural is
well documented in the catacombs of Syracuse \citep{Sgarlata2013}. It was found shattered in front of a burial, distinct
from others, known as ``the Saint's Tomb'' \citep[40-44]{Sgarlata2004}.

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img005.jpg}
\caption{360° tour of the Nassiane inscription.}
\label{fig:5}
\end{figure}

This is the Greek text:

\begin{quotation}

\begin{otherlanguage*}{greek}
chrismon ς χριστ<ι>ανῆς σεμνῆς ἀγανóφρονος [ἠδ]ὲ φιλάνδρου Nασσιανῆς τύμβον εἰσορᾷς, φίλε, κίμενον [ὣδε ἥ]τις σεμνοσύνῃσιν [ἐριζε]το Πηνελοπίῃ palma chrismon ἐνθάδε κῖτε [Nασσ]ιανή, ζήσασα [ἐν Θ(ε)ῷ] καλῶς καὶ ἀμέμπτως ἒτη λβ´, μῆνας ι´.
\end{otherlanguage*}

\end{quotation}

The inscription says: ``Oh friend, you see here the tomb of Nassiane, Christian, ripe, sweet, fond of her husband, who
competed for virtue with Penelope. Here lies Nassiane, she lived well and blamelessly in God (?), 32 years and 10
months''. This inscription documents such as epigraphic formulas affected by religious syncretism (syncretism generally
means that a complex of phenomena and concepts derive from the meeting and fusion of different religious forms) between
Pagans and Christians, for which a Christian like Nassiane could have competed in life with Penelope, the hero
Odysseus’ wife and symbol of marital devotion. Its circular shape suggested that it was a table for the ritual of
refrigerium, the funeral feast that symbolically was consumed with the dead, even this syncretic practice \ dates back
to the pagan world \citep{Sgarlata2013;Scandurra2014}.

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016BONACINIPilotprojectatPaoloOrsiMuseum-img006.jpg}
\caption{ The Nassiane inscription, recto (courtesy by the ``Paolo Orsi'' Museum). }
\label{fig:6}
\end{figure}

\section{Conclusions}

@@

Thanks to this project we hope that Google itself could realize how the time has come to ``rejuvenate'' the Google Maps
Street View system, allowing enabled users to apply additional content on the maps.

However, the wide interoperability between Google software the development of new solutions and the integration of
geo-referenced results in the page results on the search engine continues unabated: the ``Paolo Orsi'' Museum - and with
the museum, the city of Syracuse and the whole of Sicily – will surely take advantage of this new tool for its
visibility.

\section*{Acknowledgement}

To Mr. Gianfranco Guccione (\url{http://www.airworks.it/}) goes my heartfelt and deep thanks for having proposed to me this
project and having concluded it together, with the only aim of enhancing the Cultural Heritage in Sicily.

@@

I warmly thank all staff of the ``Paolo Orsi'' Regional Archaeological Museum in Syracuse for the great availability and
friendship shown to me and Mr. Guccione: Dr. Gioconda Lamagna, executive head of the Museum, Dr. Angela Maria Manenti,
head of relations with the public, all the archaeologists Dr. Anita Crispino, Dr. Agostina Musumeci and especially Dr.
Giuseppina Monterosso, tireless in supporting the project at any time and formerly member of the cabinet of Regional
Minister of Cultural Heritage and Sicilian Identity.

@@

My deepest gratitude goes to my friend Prof. Annalisa Bonica for the patience she has showed in correcting this
article.

This publication has been made on concession of the Regional Minister of Cultural Heritage and Sicilian Identity;
archaeological findings and plans presented here are exclusive property of the ``Paolo Orsi'' Regional Archaeological
Museum in Syracuse and it not allowed to be copied or duplicated by any means.

\nocite{Farman}
\bibliographystyle{sapauth-eng}
\bibliography{../../EAGLE}
